---
title: "Assignment 3: Promotions Management"
author: "Günter J. Hitsch"
date: "May 2, 2020"
output:
  pdf_document:
    number_sections: yes
    toc: yes
urlcolor: blue
graphics: yes
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = NA, message = FALSE, eval = FALSE,
                      fig.width = 4.5, fig.height = 3, fig.align = "right")
```

\setlength{\parskip}{6pt}
\newpage


```{r}
library(tidyverse)
```

\bigskip

```{r}
load("Hellmans.rdata")
```

##1.1 Creating price variable for Hellman's 32oz.mayo
```{r}
hellmans_df %>% mutate(price = dollars/units)
```

##1.2 Dividing feature and display columns by 100, summarize and comment
```{r}
hellmans_df %>% mutate (
                           feature_acv= feature_pctacv/100, 
                           display_acv = display_pctacv/100
                        ) %>%
               summarise(
                           number_of_observations = n(),
                           feature_mean = mean(feature_acv),
                           feature_median = median(feature_acv),
                           feature_sd = sd(feature_acv),
                           display_mean= mean(display_acv),
                           display_median = median(display_acv),
                           display_sd = sd(display_acv)
                        )
```


### Examination of feature and display variables

##1.3 Correlations between feature_pctacv and display_pctacv

## 1.4 Comments on findings

## 1.5 Do correlcations indicate a potential problem for your regression analysis to be performed below?


